Cheminformatics: The Computer Science of Chemical Discovery

### 2011-08-15 paper.tex
Added missing references

Graph substructure matching is a variant of \emph{graph isomorphism detection},which is
widely believed to be computationally intractable [3]. To execute a graph isomorphism
search across a full chemical database of thousands or even millions of
structures is infeasible. For that reason, complex heuristics must be
used to drastically filter the candidate search space. Structure-based
\emph{fingerprints} are commonly used for this purpose. Fingerprints encode
characteristic features of a given chemical structure in a Boolean array, or
bitmap, and they can be implemented in a variety of ways. Common fingerprints are created by an algorithm which generates bitmap patterns for each atom in the structure; then each atom and its nearest
neighbors, including the bonds between them; then each group of atoms connected
by paths of increasing lengths up to the maximum implemented for the algorithm
(commonly around eight). In this way, characteristic parts of molecule are captured and the bitmap patterns are added together into the final fingerprint. If all bits in a query fingerprint are also present in the target
fingerprint of a stored database structure, only then is the structure subjected
to the computationally expensive subgraph matching algorithm. Bit operations are
very fast and independent of the number of atoms in a structure due to the fixed
length of the fingerprint.

>>>>

Graph substructure matching is a variant of \emph{graph isomorphism
  detection},which is widely believed to be computationally
intractable \cite{cordella2001}. To execute a graph isomorphism search
across a full chemical database of thousands or even millions of
structures is infeasible. For that reason, complex heuristics must be
used to drastically filter the candidate search space. Structure-based
\emph{fingerprints} are commonly used for this purpose. Fingerprints
encode characteristic features of a given chemical structure in a
Boolean array, or bitmap, and they can be implemented in a variety of
ways. Common fingerprints are created by an algorithm which generates
bitmap patterns for each atom in the structure; then each atom and its
nearest neighbors, including the bonds between them; then each group
of atoms connected by paths of increasing lengths up to the maximum
implemented for the algorithm (commonly around eight). In this way,
characteristic parts of molecule are captured and the bitmap patterns
are added together into the final fingerprint. If all bits in a query
fingerprint are also present in the target fingerprint of a stored
database structure, only then is the structure subjected to the
computationally expensive subgraph matching algorithm. Bit operations
are very fast and independent of the number of atoms in a structure
due to the fixed length of the fingerprint.
